NGO analysis

### 2022-08-01 Dashboard/Dashboard testing material/Dashboard testing zero plot.Rmd
Zero values testing.
Excel zero plot working.

Then upload an Excel file in the set format.

@@

  ## Expenses-------------------------------------------------------------------
  expenses_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else {
      exceldata() |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          Other = expenses_other,
          Administrative = expenses_administrative,
          Fundraising = expenses_fundraising,
          `Program services` = expenses_program_services
        ) |>
        # Pivoting to use fill
        pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
        # Ordering for the plot
        mutate(
          Expenses = ordered(Expenses,
            levels =
              c("Program services", "Other", "Fundraising", "Administrative")
          )
        ) |>
        ggplot(aes(Year, value, fill = Expenses)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum())
        ) +
        scale_fill_met_d("Hokusai2") +
        labs(x = "", y = "", title = "Expenses (in thousands)")
    }
  })

@@

  ## Revenue--------------------------------------------------------------------
  revenue_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else {
      exceldata() |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          `Private contributions` = revenue_private_contributions,
          `Membership fees` = revenue_membership_fees,
          `Public grants` = revenue_public_grants,
          Investments = revenue_investments,
          `Operational activities` = revenue_operational_activities,
          `Commercial activities` = revenue_commercial_activities,
          Other = revenue_other
        ) |>
        # Pivoting to use fill
        pivot_longer(
          c(`Private contributions`:Other),
          names_to = "Revenue source"
        ) |>
        # Ordering for the plot
        mutate(
          `Revenue source` =
            ordered(`Revenue source`,
              levels =
                c(
                  "Other", "Commercial activities",
                  "Operational activities", "Investments",
                  "Membership fees", "Private contributions",
                  "Public grants"
                )
            )
        ) |>
        ggplot(aes(Year, value, fill = `Revenue source`)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum())
        ) +
        # Specifying the number of values, as the base palette only includes 6
        scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
        labs(x = "", y = "", title = "Revenue (in thousands)")
    }
  })

@@

  ## Excel data test (also above)-----------------------------------------------
  # Reactive function to test for Excel data quality (basic)
  datatest <- reactive({
    # let me know if anyone has a better idea for testing
    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }
  })
  # Let the magic/disaster happen (i.e. execute when the user clicks the button)
}) |> bindEvent(input$click)
```

@@

>>>>

Then upload an Excel file in the set format. <!--# Need to insert link when repository is public -->

@@

  ## Expenses-------------------------------------------------------------------
  expenses_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else if (excel_zero_check("Expenses") == 0) {
      zero_data()
    } else {
      exceldata() |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          Other = expenses_other,
          Administrative = expenses_administrative,
          Fundraising = expenses_fundraising,
          `Program services` = expenses_program_services
        ) |>
        # Pivoting to use fill
        pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
        # Ordering for the plot
        mutate(
          Expenses = ordered(Expenses,
            levels =
              c("Program services", "Other", "Fundraising", "Administrative")
          )
        ) |>
        ggplot(aes(Year, value, fill = Expenses)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum())
        ) +
        scale_fill_met_d("Hokusai2") +
        labs(x = "", y = "", title = "Expenses (in thousands)")
    }
  })

@@

  ## Revenue--------------------------------------------------------------------
  revenue_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else if (excel_zero_check("Revenue") == 0) {
      zero_data()
    } else {
      exceldata() |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          `Private contributions` = revenue_private_contributions,
          `Membership fees` = revenue_membership_fees,
          `Public grants` = revenue_public_grants,
          Investments = revenue_investments,
          `Operational activities` = revenue_operational_activities,
          `Commercial activities` = revenue_commercial_activities,
          Other = revenue_other
        ) |>
        # Pivoting to use fill
        pivot_longer(
          c(`Private contributions`:Other),
          names_to = "Revenue source"
        ) |>
        # Ordering for the plot
        mutate(
          `Revenue source` =
            ordered(`Revenue source`,
              levels =
                c(
                  "Other", "Commercial activities",
                  "Operational activities", "Investments",
                  "Membership fees", "Private contributions",
                  "Public grants"
                )
            )
        ) |>
        ggplot(aes(Year, value, fill = `Revenue source`)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum())
        ) +
        # Specifying the number of values, as the base palette only includes 6
        scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
        labs(x = "", y = "", title = "Revenue (in thousands)")
    }
  })

@@

  ## Excel data test (also above)-----------------------------------------------
  # Reactive function to test for Excel data quality (basic)
  datatest <- reactive({
    # let me know if anyone has a better idea for testing
    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }
  })

  ## Excel zero test------------------------------------------------------------
  excel_zero_check <- function(type) {
    excel_test() |>
      filter(
        category == as.character(all_of(type))
      ) |>
      summarise(
        zero = sum(value != 0, na.rm = TRUE)
      ) |>
      pull()
  }

  # Let the magic/disaster happen (i.e. execute when the user clicks the button)
}) |> bindEvent(input$click)
```

@@

## zero_data_plot --------------------------------------------------------------

zero_data <- function() {
  empty_plot("All values are zero.\nThis is unusual and may\nindicate a data problem.")
}
